---
title: "R - Pumpkins Output"
format: html
editor: visual
---

## Pumpkins Dataset Analysis

### 1. Read in pumpkins.csv data
```{r}
#| include: false
#| echo: false
pumpkins <- read.csv("C:/Users/hanna/OneDrive/Documents/Computational Bio/Coding Challenge/pumpkins_03.csv")
library(tidyverse)
library(gt)
```
### 2. Identify the heaviest pumpkin based on variety, location and year
```{r}
#| echo: false
pumpkins_2 <- pumpkins %>%
  mutate(location=paste(city, state_prov, country, sep = ", ")) 
pumpkins_heavy <- pumpkins_2 %>%
  select(variety, location, id, weight_lbs) %>%
  slice_max(weight_lbs)
gt(pumpkins_heavy)
 
```
### 3. Create a new column "weight_kg" based on pounds(lbs) -> kilograms(kg) conversion function

```{r}
#| echo: false
pounds_to_kilos <- (kilos = pumpkins_2$weight_lbs * 0.453592)
pumpkins_3 <- pumpkins_2 %>%
    mutate(weight_kg=pounds_to_kilos, .after=weight_lbs)
pumpkins_3 %>%
  head(10) %>%
  gt() %>%
  tab_header(
     title = "Pumpkins Dataset",
     subtitle = "Including new column 'weight_kg'"
  )
```
### 4. Create a new column "weight_class"
First find the maximum, minimum and mean weights (lbs) to determine class parameters
```{r}
#| echo: false
pump_summ <- pumpkins_3 %>%
    summarise(mean_weight = mean(weight_lbs, na.rm = TRUE),
              max_weight = max(weight_lbs, na.rm = TRUE),
              min_weight = min(weight_lbs, na.rm = TRUE))
gt(pump_summ)
```
The weights(lbs) of the pumpkins have been categorised into "Light" (0-750 lbs), "Medium" (750-1500 lbs) and "Heavy" (>1500 lbs)
```{r}
#| echo: false
pumpkins_4 <- subset(pumpkins_3, est_weight != 0.0000 & est_weight < 9000) 
pumpkins_wc <- pumpkins_4 %>%
    mutate(weight_class = case_when(
      weight_lbs <= 750 ~ 'Light (<750 lbs)',
      weight_lbs <=1750 ~ 'Medium (750-1500 lbs)',
      TRUE ~ 'Heavy (>1500 lbs)'
    ))
pumpkins_wc %>%
    head(10) %>%
  gt() %>%
  tab_header(
     title = "Sample of Pumpkins Dataset",
     subtitle = "Including new column 'weight_class'"
  )
```

### 5. Estimated vs. actual weight scatter plot
Values with and estimated weight of "0.0000" have been removed as it is assumed the value was given instead of "NA"
Values with an estimated weight of < 9000 lbs have been removed as it is assumed they're a result of input error
```{r}
#| echo: false
est_vs_actual_plot <- ggplot(pumpkins_wc, aes(x = est_weight, y = weight_lbs, colour=weight_class))+
  geom_point(size = 0.5)+
  scale_colour_manual(values = c('Light (<750 lbs)' = 'hotpink2',
                               'Medium (750-1500 lbs)' = 'deeppink2',
                               'Heavy (>1500 lbs)' = 'deeppink4'))+
  labs(title = "Estimated vs. Actual Weight (lbs) Plot", x = "Estimated Weight (lbs)", y = "Actual Weight (lbs)", colour = "Weight Class") 
est_vs_actual_plot
```
### 6. Filter data to only include pumpkins from three countries
First determined the countries with the most pumpkin entries to decide which three countries to select
```{r}
#| echo: false
pumpkins_countries <- pumpkins %>%
  count(country) %>%
  arrange(desc(n))
pumpkins_countries %>%
  head(5) %>%
  gt()
```
From this Canada, Germany and United States were selected
```{r}
#| echo: false
pumpkins_filtered <- pumpkins %>%
  filter(country == "United States" | country == "Canada" | country == "Germany") 
pumpkins_filtered%>%
  head(10) %>%
  gt() %>%
  tab_header(
    title = "Filtered Pumpkins Dataset",
    subtitle = "Top 3 Countries with The Most Pumpkin Entries "
  )
```
This filtered data is then saved and available as "pumpkins_filtered.csv"

### 7. From the "pumpkins_filtered":
#### a. Identify the mean weight of pumpkins for each of the three countries. Which had the highest mean pumpkin weight?
```{r}
#| echo: false
pumpkins_filtered_country <- pumpkins_filtered %>%
  group_by(country) %>%
  summarise(mean_weight_lbs = mean(weight_lbs, na.rm = TRUE))
pumpkins_max_weight <- pumpkins_filtered_country %>%
  slice_max(mean_weight_lbs)

pumpkins_max_weight %>%
  head() %>%
  gt()
```

#### b. Identify the mean weight for each variety of pumpkin for each country. Which country had the lowest mean weight?
```{r}
#| echo: false
#| warning: false
pumpkins_filtered2 <- pumpkins_filtered %>%
  group_by(country, variety) %>%
  summarise(mean_weight_lbs = mean(weight_lbs, na.rm = TRUE))

pumpkins_min_weight <- pumpkins_filtered2 %>%
  slice_min(mean_weight_lbs)

pumpkins_min_weight %>%
  head() %>%
  gt()
```
### 8. From "pumpkins_filtered" plot the pumpkin weight distributions for the three countries as a boxplot
```{r}
#| echo: false
filtered_weight_boxplot <- ggplot(aes(x = country,y = weight_lbs, fill = country), data = pumpkins_filtered)+
  geom_boxplot()+
  labs(x = "Country", y = "Weight (lbs)", title = "Boxplot of Weight Distribution per Country")+
  scale_fill_manual(values = c('Canada' = 'hotpink2',
                               'Germany' = 'olivedrab3',
                               'United States' = 'steelblue3'))

filtered_weight_boxplot
```
### 9. From "pumpkins_filtered" plot a facet plot showing the data from each variety of pumpkin as a separate sub-plot
```{r}
#| echo: false
facet_plot <- ggplot(aes(country, weight_lbs, fill = country), data = pumpkins_filtered)+
  geom_boxplot()+
  facet_wrap(~variety)+
  labs(x = "Country", y = "Weight (lbs)", title = "Facet Plot of Weight Distribution per Country for each Variety")+
  theme(
    axis.text.x = element_text(angle = 90)
  )+
  scale_fill_manual(values = c('Canada' = 'hotpink2',
                               'Germany' = 'olivedrab3',
                               'United States' = 'steelblue3'))

facet_plot
```


